---
title: "insurance regression"
author: "Zhivko Dimitrov"
date: "21 March 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

importing my data

```{r}
claims <- read.csv('data/claims.csv')
```

loading dplyr

```{r}
library(dplyr)
```

i want to convert my data to a tbl so i can start exploring it

```{r}
tbl_df(claims)
glimpse(claims)
```

for the purposes of my study  i am only interested in a few of these columns, so i will clean up my dataset to remove all data i do not need:

```{r}
claims <- select(claims, State, Claim.Reason, Claim.Amount )
glimpse(claims)
```

i now want to summarize the number of cases of hail claims per state, their sums and their means

```{r}
claims <- claims %>%
  group_by(State) %>%
  filter(Claim.Reason=="Hail") %>%
  summarise(NumberClaims=n(), SumAllClaims=sum(Claim.Amount), MeanClaim=mean(Claim.Amount))
```

now i will import and process the data for hail
```{r}
hail <- read.csv('data/hail.csv')
tbl_df(hail)
glimpse(hail)
```
i am not intersted in the data and x columns, as well as in all cases of hail that did not affect any population or data for states that are not included in the claims dataset


```{r}
hail <- select(hail, State, Hail.Size, Population.Affected)
hail <- filter(hail, Population.Affected > 0, State ==c("Iowa","Kansas","Missouri","Nebraska","Oklahoma"))
```

i see that the hail size variable is not a number and will need to be normalized. lets see all the possible values

```{r}
unique(hail$Hail.Size)
```
i will need to extract the numbers from these values

```{r}
library(tidyr)
hail <- hail %>%
  mutate_each(funs(extract_numeric), Hail.Size)
```

now i want to compute the weigted average size of hail for each state and also the total population affected. i also hypothesize that bigger hail size might have a  disproportionate impact on insurance claims so i woukd also want to test against the maximum hail size too.

```{r}
hail <- hail %>%
  group_by(State)%>%
  summarise(TotalPopulationAffected = sum(Population.Affected), AverageHailSize = weighted.mean(Hail.Size, Population.Affected), MaxHailSize = max(Hail.Size))

```

now finally i want to join the claims and the hail datasets

```{r}
ClaimsHail <- left_join(claims, hail, by="State")

```









